---
title: "Natasha Woon"
output: github_document
---

This file is for doing data imports 

```{r}
library(tidyverse)
library(readxl)
library(haven)
```

Import our first dataset

```{r}
litters_df = 
  read_csv(
    file = "data_import_examples/FAS_litters.csv",
           na = c("", "NA", ".")
           ) 

#skip allows us to skip rows in the csv file

litters_df = janitor::clean_names(litters_df)
```

Import our second dataset 

```{r}
pups_df=
  read_csv(
    file = "data_import_examples/FAS_pups.csv", 
    skip = 3,
    na = c("", "NA", ".")
    ) 

pups_df = janitor::clean_names(pups_df)
```

## Look at the data 

```{r}
litters_df 
pups_df
```

Head or tail 

```{r}
head(litters_df, 20) #this shows the first 20 rows

tail(pups_df, 5) # this shows the last 5 rows
```

skimming 

```{r}
skimr::skim(pups_df) # prints an extended summary of what's going on #able to look at what is missing, what is complete, the mean, sd, outlines of a histogram
```


## oh excel ... 

Jenny Bryan made 'readxl' solve our problems. 

```{r}
mlb_df = read_excel("data_import_examples/mlb11.xlsx")
```

Look at the data!
```{r}
mlb_df
```

Load some LotR data! 

Import FOTR words

```{r}
fotr_df = read_excel(
  "data_import_examples/LotR_Words.xlsx",
  range = "B3:D6"
  )

fotr_df
```

What about two towers?

```{r}
tt_df = read_excel(
  "data_import_examples/LotR_Words.xlsx",
  range = "F3:H6"
  )

tt_df
```

## Import SAS 

Read in the PULSE dataset

```{r}
pulse_df = 
  read_sas("data_import_examples/public_pulse_data.sas7bdat")

pulse_df = janitor::clean_names(pulse_df)
```


## What's so bad about `read.csv()`

```{r}
pups_df_based = 
  read.csv("data_import_examples/FAS_pups.csv") ##output many many many rows so it's a bit tacky

pups_df_based
```

Won't show: lazy about variable names; super slow for a big dataset. 